### 2022-08-30 Reports/Week5/weekly_report5.tex
weekly report 5

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makecell}
\usepackage[margin=1in]{geometry}

\author{Lukas Evers}
\title{Weekly Report 3}

\renewcommand\theadalign{bl}
\renewcommand{\cellalign}{tc}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}

\begin{document}

\maketitle
\newpage

\section{Scenarios}

\subsection{Scenarios with Subscenarios}

\begin{table}[!h]

	\caption{Sense-Plan-Act Scenarios}
	\label{tab:table2}
	\begin{tabular}{|c|l|l|}
		\hline
      	\thead{Scenario Name} & \thead{Sub Scenario} & \thead{Description} \\
      	\hline
      	LidarCrash_1 &  \makecell{Lidar Node dies while stationary} & \makecell{Crash of the Lidar node.} \\
      	\hline
      	LidarCrash_2 &  \makecell{Lidar Node dies while driving straight} & \makecell{Crash of the Lidar node.} \\
      	\hline
      	LidarCrash_3 &  \makecell{Lidar Node dies while driving around a corner} & \makecell{Crash of the Lidar node.} \\
     	\hline
      	OdomCrash_1 &  \makecell{Odometry Node dies while standing} & \makecell{Crash of the Odometry.} \\
      	\hline
      	OdomCrash_2 &  \makecell{Odometry Node dies while driving} & \makecell{Crash of the Odometry.} \\
      	\hline
      	IMUCrash_1 &  \makecell{Odometry Node dies while standing} & \makecell{Crash of the IMU Driver.} \\
      	\hline
      	IMUCrash_2 &  \makecell{Odometry Node dies while driving} & \makecell{Crash of the IMU Driver.} \\
      	\hline
      	NoPathFound_1 & \makecell{No valid global path can be found}  & \makecell{Path can not be calculated.} \\
		\hline      	
      	NoPathFound_2 & \makecell{Global Planner Node crashed.}  & \makecell{No response from Path Planner.} \\
      	\hline
      	ForcedCollision_1 & \makecell{Obstacle crashes into robot}  & \makecell{The robot is following its path and something else crashes into the robot.} \\
      	\hline
      	ForcedCollision_2 & \makecell{The robot crashes into undetected obstacle}  & \makecell{The robot is following its path crashes into an .} \\
      	\hline 
      	BatteryLow_1 & \makecell{Battery Voltage Low} & \makecell{The remaining battery is too low to reach the destination.} \\
      	\hline
      	MotorFailure_1 & \makecell{The motor driver fails.} & \makecell{The motor has a hardware failure.} \\
      	\hline
      	NetworkFailure_1 & \makecell{The robot can not receive any commands.} & \makecell{The method to transmit the motor commands crashes.} \\
      	\hline

	\end{tabular}
\end{table}

\newpage

\section{Next Week}

\end{document}