### 2022-09-06 Thesis/text/ergebnisse.tex

\chapter{Evaluation}
\label{cha:ergebnisse}

\section{Scenarios}